### 2024-10-30 progress/progress.tex
readme; progress template

\documentclass[12pt]{article}

\usepackage[margin = 1in]{geometry}
\usepackage{authblk}
\usepackage[inline]{enumitem}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{setspace}
\onehalfspacing

\begin{document}

\title{Progress Report: Transformer-CRF Integration for Sequence Labeling on
EEG Data}

\author[1]{Xiaohang Ma}
\author[2]{Shiying Xiao}
\author[2]{Xiaohui Yin}

\affil[1]{Department of Mathematics, University of Connecticut}
\affil[2]{Department of Statistics, University of Connecticut}

\date{\today}

\maketitle

\section{Progress}
% Overview of progress made and key accomplishments.

\section{Preliminary Results}
% Summary of initial findings and analyses.

\section{Challenges}
% Discussion of any challenges faced during the project.

\section{Plan Changes}
% Outline of any adjustments made to the original plan.

\section{Next Steps}
% Description of remaining tasks and strategy for completion.

\section{Updates}
% Additional relevant updates or future directions.

\end{document}