### 2024-11-09 manuscript/main.tex
add ELBO

\documentclass[conference]{IEEEtran}

%% keep it clean; only include those you need
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{bbm}
\usepackage{graphicx}
\graphicspath{{./}{../image/}}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{xcolor}

%% notations; only define those you use frequently
\newcommand{\EE}{{\mathbb{E}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Ib}{\mathbf{I}}
\newcommand{\hb}{\mathbf{h}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Ocal}{\mathcal{O}}

\begin{document}

\title{\LARGE Transformer-CRF Integration for Sequence Labeling on EEG Data}

\author{\IEEEauthorblockN{Xiaohang Ma}
\IEEEauthorblockA{\textit{Department of Mathematics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohang.ma@uconn.edu}
\and
\IEEEauthorblockN{Shiying Xiao}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
shiying.xiao@uconn.edu}
\and
\IEEEauthorblockN{Xiaohui Yin}
\IEEEauthorblockA{\textit{Department of Statistics} \\
\textit{University of Connecticut}\\
Storrs, CT, USA \\
xiaohui.yin@uconn.edu}
}

\maketitle

\begin{abstract}

In this project, we will integrate powerful deep neural networks and
statistical methods with a probabilistic graphical model (PGM) to effectively
tackle sequence labeling tasks on a heterogeneous medical care dataset.

\end{abstract}

\begin{IEEEkeywords}

probabilistic graphical model

\end{IEEEkeywords}

\section{CNN-HCRF Model}

The classical hidden conditional random field (HCRF) model can be formulated as
\begin{equation}
\begin{split}
\PP(y \vert \xb; \theta) &= \sum_{\hb \in H^N}
\mathbb{P}(y, \hb \vert \xb; \theta) \\
&= \frac{\sum_{\hb \in H^N} \exp\left(\Phi(y, \hb, \xb; \theta)\right)}
{\sum_{y^\prime} \sum_{\hb \in H^N}
\exp\left(\Phi(y^\prime, \hb, \xb; \theta)\right)}.
\end{split}
\end{equation}

Different to the design of most existing HCRF models, which they assign
potentials based on the features predefined from the input data $\Ib$.
%$\PP(y, \hb \vert \xb; \theta)$.
%Here, we consider a generative model that constructing
%$\PP(\hb \vert \xb; \theta)$ through a CRF framework then using an 
%unary potential for , 
Our HCRF replace $I$ with features learned by CNN, i.e. $\Fb(\Ib)$. 
Thus the new potentials become:
\begin{equation*}
\Phi(y, \hb, \xb; \theta) = \Phi(y, \hb, \Fb(\Ib); \theta).
%\PP(y, \hb \vert \xb; \theta) = \PP(y \vert \hb, \xb; \theta)
%\cdot \PP(\hb \vert \xb; \theta).
\end{equation*}
Thus, we design the potential of our HCRF model as.
\begin{equation}
\begin{split}
\Phi(y, \hb, \xb; \theta) &= \underbrace{
\sum_{j \in \nu} \phi(h_j, x_j; \omega)
+ \sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
% \propto \log\PP(\hb \vert \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}} \\
&+ \underbrace{
\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)}_{
% \propto \log\PP(y | \hb,  \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}}
\end{split}
\end{equation}

{\textbf{Unary Potential}}
$\phi(h_j, x_j; \omega)$ measures the likelihood of the local feature
$x_j$ is assigned as the hidden attention state $h_j$.
We generate the likelihood through a softmax operation on $\Fb(\xb)$
the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
by which the potential of each pixel $x_j$ assigned as each state of
hidden state set $\{0, 1, \dots, 5\}$ is drawn and represented by
a matrix with $(N, 2)$ dimension.

The parameter $\omega$ is learned by our end-to-end CNN-Unet training 
structure.

{\textbf{Unary Potential}}
$\varphi(y, h_j; \delta)$ measures the compatibility between global
class label $y$ and the hidden local state $h_j$. This potential is
parametrized as:
\begin{equation*}
\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
\cdot \one(y = a) \cdot \one(h_j = b).
\end{equation*}

{\textbf{Binary Potential}}
$\psi(h_i, h_j, x_i, x_j; \eta)$
balances the hidden label compatibility of the neighboring pixels on the 
feature image $\Fb(\Ib)$.
We define this potential through a common contrast-sensitive two-kernal 
potentials~\cite{krahenbuhl2011efficient}\cite{chen2022end}:
\begin{equation}
\begin{split}
& \psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
\omega_1 \exp \bigg(
-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2} \\
&\quad
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib)\right\rvert^2}{2\eta_\beta^2}
\bigg)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
\right)
\Bigg]
\end{split}
\end{equation}

\section{Variational Inference of HCRF}

\subsection{ELBO Under Mean-Filed Variational Family}

The existence of the hidden variable $\hb$ makes our model more interpretable.
However, since We will train our model through maximal likelihood method,
introducing the hidden variable will make precisely computation the summation
$\sum_{\hb \in H^N} \PP(y, \hb \vert \xb; \theta)$ intractable.
To address this issue, we make our computation more feasible through a
mean-field variational inference framework, which enables our model
efficiently estimating the log-likelihood $\log\PP(y \vert \xb; \theta)$.

Firstly, we introduce the mean-field varepsilon family approximation of
$\log\PP(\hb \vert y, \xb; \theta)$ as:
\begin{equation*}
\QQ(\hb) = \prod_{i=1}^N q_i(h_i).
\end{equation*}
Then the evidence lower bound (ELBO) under the mean-field family is given by
\begin{equation}
\begin{split}
& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) \\
&\quad = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right] \\
&\quad + \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}} 
\QQ(h_i = l) \QQ(h_j = l^\prime) \mu(l, l^\prime) \\
&\quad \Bigg[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right) \\
&\quad + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \Bigg] \\
&\quad - \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
\end{split}
\end{equation}

\subsection{Updating Variational Family}

By the coordinate ascent variational inference(CAVI) algorithm,
we can obtain that fix $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$
that maximizes $\textrm{ELBO}(q_i(h_i))$ is given by
\begin{equation}
\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}.
\end{equation}
Thus, we can get that
\begin{equation}
\begin{split}
q_i(l) &= \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \\
&+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) 
\Bigg[ \omega_1 \exp \big(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}\\
&- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\big)
+ \omega_2 \exp \big(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\big)
\Bigg]
\Bigg\}
\end{split}
\end{equation}

\subsection{Computational Complexity Analysis}

\begin{itemize}
\item {\textbf{Updating Variational Family}} $\{q_i(h_i)\}$:
The computationally expensive part of the fixed point iteration of $q_i(h_i)$
comes from the message passing, i.e. the convolution of $\QQ(\hb)$ and
Gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$. We need $\Ocal(N^2)$
runtime for $N$ pixels when evaluating this term precisely.
To make the computation more feasible, a truncated gaussian kernal will be
employed, which is only supported on spaces within certain proportion of
the untruncated standard deviation. Then the approximate message passing is
linear in the number of pixels $N$.
\item {\textbf{Computation of the ELBO}}: The summation over the product of
$q_i(h_i)q_j(h_j)$ and the gaussian kernal in $\psi(h_i, h_j, x_i, x_j; \eta)$
also requires a $\Ocal(N^2)$ complexity in time.
Owes to a similar truncated gaussian kernal approxiamtion, the runtime can
be reduced to $\Ocal(N)$.
\end{itemize}

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}