### 2024-11-11 progress/progress.tex
reformat report

\documentclass[10pt]{article}

\usepackage[left=.5in, right=.5in, top=.3in, bottom=.3in]{geometry}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{authblk}
%% put affiliations into one line
\makeatletter
\renewcommand\AB@affilsepx{\quad\protect\Affilfont}
\usepackage{bbm}
\usepackage[inline]{enumitem}
\usepackage{graphicx}
\graphicspath{{./}{../image/}}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{setspace}
\setstretch{1.2}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{8pt}{8pt}

\newcommand{\EE}{{\mathbb{E}}}
\newcommand{\PP}{{\mathbb{P}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\Fb}{\mathbf{F}}
\newcommand{\Ib}{\mathbf{I}}
\newcommand{\hb}{\mathbf{h}}
\newcommand{\xb}{\mathbf{x}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Ocal}{\mathcal{O}}

\begin{document}

\title{\vspace{-1cm} \Large
Progress Report: Transformer-CRF Integration for Sequence Labeling on EEG Data}

\author[1]{Xiaohang Ma}
\author[2]{Shiying Xiao}
\author[2]{Xiaohui Yin}

\affil[1]{Department of Mathematics, UConn}
\affil[2]{Department of Statistics, UConn}

\date{\vspace{-1.3cm}}

\maketitle

\section{Progress}
% Overview of progress made and key accomplishments.

The project is progressing according to schedule. To date, we have made
significant strides in several key areas:
\begin{enumerate}
\item Model Development:
We have initiated development of a convolutional neural network (CNN) combined
with a hidden conditional random field (HCRF). This hybrid approach aims to
create a robust image segmentation framework capable of capturing both local
and global dependencies within the data structure.
\item Variational Inference Implementation:
To efficiently approximate the posterior distributions in our complex model,
we have integrated variational inference into the HCRF component.
This technique is particularly effective in handling the computational
challenges posed by the dataset's heterogeneity.
\item Data Preprocessing:
We completed the data preprocessing for EEG signals and their corresponding
spectrograms. The EEG signals were transformed into spectrogram images using
built-in functions from open-source libraries such as OpenCV and PyWavelets.
We then concatenated these spectrograms into a 512 $\times$ 512 image,
which is now ready to be used as input for our model.
\end{enumerate}
%This progress has provided a foundation for the next
%steps of our research.

\section{Preliminary Results}
% Summary of initial findings and analyses.

In our model, the variable $y$ represents the corresponding brain signal
label of the inpute image $\xb$. The hidden variable $\hb= \{h_i\}$
represents the label at each pixels of $\xb$. The parameter $\theta$
contain parameters of neural networks as well as the HCRF.
We introduce the potential functions of our HCRF model as:
\begin{equation*}
\Phi(y, \hb, \xb; \theta) = \underbrace{
\sum_{j \in \nu} \phi(h_j, x_j; \omega) +
\sum_{i \neq j} \psi(h_i, h_j, x_i, x_j; \eta)}_{
% \propto \log\PP( \hb \vert \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(\hb \vert \xb; \theta)$}}
+ \underbrace{
\sum_{j \in \nu} \varphi(y, h_j, x_j; \delta) + \vartheta(y, \xb; \varpi)
}_{
% \propto \log\PP(y | \hb, \xb; \theta)
\textrm{Measures log-likelihood $\log\PP(y \vert \hb, \xb; \theta)$}
}
\end{equation*}

\begin{itemize}
\item \textbf{Unary Potential}
$\phi(h_j, x_j; \omega)$ measures the likelihood of the local feature
$x_j$ is assigned as the hidden attention state $h_j$.
We generate the likelihood through a softmax operation on $\Fb(\xb)$
the feature maps of CNN with $(N, C_{\textrm{out}})$ dimensions,
by which the potential of each pixel $x_j$ assigned as each state of
hidden state set $\{0, 1, \dots, 5\}$ is drawn and represented by
a matrix with $(N, 2)$ dimension.
\item The parameter $\omega$ is learned by our end-to-end CNN-Unet
training structure.
\item \textbf{Unary Potential}
$\varphi(y, h_j; \delta)$ measures the compatibility between global class
label $y$ and the hidden local state $h_j$. This potential is parametrized as:
\begin{align*}
\varphi(y, h_{j}; \delta) = \sum_{a \in Y} \sum_{b \in H} \delta_{a, b}
\cdot \one(y = a) \cdot \one(h_j = b).
\end{align*}
\item \textbf{Binary Potential}
$\psi(h_i, h_j, x_i, x_j; \eta)$ balances the hidden label compatibility
of the neighboring pixels on the feature image $\Fb(\Ib)$.
We define this potential through a common contrast-sensitive two-kernal 
potentials~\citep{krahenbuhl2011efficient,chen2022end}:
\begin{equation*}
\psi(h_i, h_j, x_i, x_j; \eta) = \mu(h_i, h_j) \Bigg[
\omega_1 \exp \left(
-\frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
-\frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2} 
\right) + \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2 \eta_\gamma^2}
\right)
\Bigg]
\end{equation*}
\end{itemize}

The mean-field varepsilon family approximation of
$\log\PP(\hb \vert y, \xb; \theta)$ as $\QQ(\hb) = \prod_{i=1}^N q_i(h_i)$.
Then the evidence lower bound under the mean-field family is given by
\vspace{-1cm}
\begin{equation*}
\begin{split}
& \textrm{ELBO}(\QQ) = \EE_{\QQ(\hb)} \log\PP(y, \hb \vert \xb; \theta)
- \EE_{\QQ(\hb)} \log q(\hb) = \sum_{i=1}^N \EE_{q_i(h_i)}
\left[ \phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta) \right]
+ \sum_{i=1}^N \sum_{j=1}^N \sum_{l,l^\prime} \one_{\{i \neq j\}}\QQ(h_i = l) \\
&\quad \QQ(h_j = l^\prime) \mu(l, l^\prime)
\Bigg[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\gamma^2}
\right) \Bigg]
- \sum_{i=1}^N \EE_{q_i} \log q_i(h_i) + \vartheta(y, \xb; \varpi).
\end{split}
\end{equation*}

By the coordinate ascent variational inference(CAVI) algorithm,
we can obtain that fix $q_j(h_j), \forall j \neq i$, the optimal $q_i(h_i)$
that maximizes $\textrm{ELBO}(q_i(h_i))$ is given by
$\QQ(h_i) \propto \exp\{\EE_{\QQ(\hb_{-i})}
\log\PP(y, h_i, \hb_{-i} \vert \xb; \theta)\}$.
Thus, we can get that
\begin{equation*}
\begin{split}
q_i(l) &= \frac{1}{Z_i} \exp \Bigg\{
\phi(h_j, x_j; \omega) + \varphi(y, h_j, x_j; \delta)
+ \sum_{j \neq i} \sum_{l^\prime} q(l^\prime) \mu(l, l^\prime) \\
&\quad \left[ \omega_1 \exp \left(
- \frac{\left\lvert p_i - p_j \right\rvert^2}{2\eta_\alpha^2}
- \frac{\left\lvert \Fb_i(\Ib) - \Fb_j(\Ib) \right\rvert^2}{2\eta_\beta^2}
\right)
+ \omega_2 \exp \left(
- \frac{\left\lvert p_i - p_j 
\right\rvert^2}{2\eta_\gamma^2}
\right)
\right]
\Bigg\}
\end{split}
\end{equation*}

After building the theoretical model formualtion of our project, we also
finished preliminary analyses of the data. Our preprocessed spectrograms
dataset that are ready for training model. The dataset will be available on
\href{https://github.com/Carol-seven/Transformer-CRF_EEG}{GitHub} and
made public after the project is completed.

\section{Challenges}
% Discussion of any challenges faced during the project.

We encountered minor challenges with data handling and computational
limitations. The complexity of certain calculations required
additional processing time, slightly delayed our progress.
We are exploring ways to optimize these computations.

\section{Plan Changes}
% Outline of any adjustments made to the original plan.

\begin{itemize}
\item \textbf{CRF Layer Modification:}
Our motivation for introducing the CRF layer was to allow the model to
identify regions of input images that correspond to abnormal brain activity
signals. However, the original dataset, including spectrograms and
EEG-transformed spectrograms, lacked segmentation information, meaning no
labels were available to train the CRF.
Our initial solution was to generate pixel-wise CRF labels based on
the global ground truth distribution. However, further review revealed that
this approach could lead to isolated label regions, which would counteract
our intent for CRF integration. Consequently, we modified the CRF layer into
a HCRF by setting pixel labels as hidden variables and adopted a variational
inference framework to manage these hidden variables in the CRF model.
\item \textbf{Model Structure Adjustment:}
We replaced the Transformer-based architecture with a U-Net CNN structure.
This choice reflects findings in recent research that fine-tuning large models
like Transformers can be time-consuming. Instead, we opted for the
well-established U-Net CNN, a classical model for image segmentation,
which integrates effectively with the CRF layer.
\end{itemize}

\section{Next Steps}
% Description of remaining tasks and strategy for completion.

The remaining tasks include:
\begin{enumerate*}[label = (\roman*)]
\item Expanding and refining the visualizations based on updated data.
\item Finalizing the construction of the neural network model, including
the U-Net CNN layer and HCRF layer, based on the formulas we have derived.
\item Training and tuning the model on our preprocessed data.
\item Conducting a comprehensive analysis with the refined model.
\item Preparing the final report and presentation materials.
\end{enumerate*}

\section{Updates}
% Additional relevant updates or future directions.

We will provide regular updates on model performance, challenges,
and any significant changes to the project plan.

\footnotesize
\bibliography{../manuscript/refs}
\bibliographystyle{chicago}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: